
#load packages needed
```{r}
library(CellMembrane)
library(RIRA) 
library(Seurat)
library(Rdiscvr)
library(Rlabkey)
library(UCell)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(tibble)
library(ggpubr)
library(enrichR)
library(readxl)
library(forcats)
library(ggpubr)
library(ggstatsplot)
setEnrichrSite("Enrichr") # Human genes
dbs <- listEnrichrDbs()
dbs_go <- c("GO_Biological_Process_2021")
dbs_pathway <- c("MSigDB_Hallmark_2020", "BioCarta_2016", "Reactome_2022", "DSigDB", "WikiPathways_2019_Human", "KEGG_2021_Human", "DisGeNET", "WikiPathway_2021_Human", "HDSigDB_Human_2021")
```

#Load necessary functions
```{r}
#Load necessary functions for pseudobulk
RunPairwiseContrasts <- function(fit, test.use, logFC_threshold = 1){
  #create a nx2 array of all possible unique pairwise combinations of contrasts
  contrasts <- t(utils::combn(colnames(fit$design), m = 2))
  results <- future.apply::future_lapply(split(contrasts, 1:nrow(contrasts)) , FUN = function(x){
    up_contrast<- x[[1]]
    down_contrast <- x[[2]]
    contrast_name <- paste0(up_contrast, "-", down_contrast)
    print(paste("Performing DE for: ", contrast_name))
    contrast <- limma::makeContrasts(contrasts = contrast_name, levels = fit$design)
    result <- PerformDifferentialExpression(fit, contrast, contrast_name, logFC_threshold = logFC_threshold, FDR_threshold = 0.05, test.use = test.use)
  })
  #use contrast names to label the results list
  names(results) <- paste0(contrasts[,1],  "-", contrasts[,2])
  return(results)
}

#Load necessary TCR functions
Plot_bar_topvalues <- function(so = NULL, feat = NULL, title="", thr = 1){
  if(is.null(so)) print("give a seurat object to so")
  if(is.null(feat)) print("give a meta feature to tabulate to feat")
  print(feat)
  
  TCR_tempVec <- so@meta.data[, feat]

  TCR_tempVec_tab <- sort(table(TCR_tempVec), decreasing = T)

  TCR_tempVec_tab_gt1 = TCR_tempVec_tab[TCR_tempVec_tab>thr]

  ggplot(reshape2::melt(TCR_tempVec_tab_gt1) , aes(TCR_tempVec, value)) +
    ggplot2::geom_bar(stat="identity") + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    ggplot2::ggtitle(paste0(title, "\nlen=",length(TCR_tempVec_tab_gt1) )) 
}

plot_bar_pairtopvalues_facetted <-function(so = NULL, feat1 = NULL, feat2 = NULL, title="", thr = 1){
  if(is.null(so)) print("give a seurat object to so")
  if(is.null(feat1)) print("give a meta feature to tabulate to feat")
  if(is.null(feat2)) print("give a meta feature to tabulate to feat")
  
  PairWiseDF <- subset(data.frame(table(so@meta.data[, feat1], so@meta.data[,feat2])), Freq > thr)

# need to log since the difference betwwen the biggest and lowest counts is huge

#facetted 
 ggplot(PairWiseDF , aes(Var1, log10(Freq), fill=Var2)) +
    ggplot2::geom_bar(stat="identity") + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    ggplot2::ggtitle(paste0(title, "\nlen=",nrow(PairWiseDF) )) + facet_wrap(~Var2)
 

}

plot_bar_pairtopvalues_stacked <-function(so = NULL, feat1 = NULL, feat2 = NULL, title="", thr = 1){
  if(is.null(so)) print("give a seurat object to so")
  if(is.null(feat1)) print("give a meta feature to tabulate to feat")
  if(is.null(feat2)) print("give a meta feature to tabulate to feat")
  
  PairWiseDF <- subset(data.frame(table(so@meta.data[, feat1], so@meta.data[,feat2])), Freq > thr)

# need to log since the difference betwwen the biggest and lowest counts is huge

#singular and stacked
  ggplot(PairWiseDF , aes(Var1, log10(Freq), fill=Var2)) +
    ggplot2::geom_bar(stat="identity", position="stack") + ggplot2::theme_bw() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) + 
    ggplot2::ggtitle(paste0(title, "\nlen=",nrow(PairWiseDF) )) 
  
}
```

#Load data files if changes
```{r}
#saveRDS(seuratObj, file = "./CB_paper_seurat.rds")
#saveRDS(seuratObjTRAVpos, file = "./CB_paper_TRAVpos.rds")
#saveRDS(seuratObjTRAVneg, file = "./CB_paper_TRAVneg.rds")
#saveRDS(seuratObj@meta.data, file = "./metadata.rds")
```

#Load data files
```{r}
fn <- 'CB_paper_seurat.rds'
seuratObj <- readRDS(fn)

fn <- 'CB_paper_TRAVpos.rds'
seuratObjTRAVpos <- readRDS(fn)

fn <- 'CB_paper_TRAVneg.rds'
seuratObjTRAVneg <- readRDS(fn)

meta <- readRDS("./metadata.rds")
```

_____________
#Figure 1a-c
```{r, echo = TRUE, fig.show='asis'}
a1 <- DimPlot(seuratObj, group.by = "Donor", cols = c("dark red", "dark blue")) +
  ggtitle("")  +  guides(color = guide_legend(override.aes = list(size=10), ncol=1)) +
  theme(legend.text = element_text(size = 20)) + 
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))

b1 <- DimPlot(seuratObj, group.by = "ClusterNames_0.4", label = T, label.size = 8) + 
  ggtitle("") + 
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.text = element_text(size = 20),
        axis.line = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))

df <- data.frame(table(seuratObj$ClusterNames_0.4, seuratObj$Donor))
colnames(df) = c("Cluster", "Donor Age", "Count")

df1 <- data.frame(table(seuratObj$Donor))
colnames(df1) = c("Donor Age", "Total")

df2 <- left_join(df, df1, by = "Donor Age")
df2$Proportion <- df2$Count/df2$Total

c1 <- df2 %>% 
  ggplot(aes(x = Cluster, 
             y = Proportion, 
             fill = `Donor Age`)) + 
  geom_bar(position = "fill", stat = "identity") + 
  scale_fill_manual(values = c("dark red", "dark blue")) +
  theme(legend.text = element_text(size = 20), 
        axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20), 
        axis.text.y = element_text(size = 14), 
        axis.text.x = element_text(size = 14), 
        legend.title = element_blank())

a1
b1
c1
```

#Figure 1d
```{r, echo = TRUE, fig.show='asis'}
d1a <- FeaturePlot(seuratObj, features = c("adt_CD4"), max.cutoff = 'q95', min.cutoff = 'q05') +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("CD4 Prt Exp")

d1b <- FeaturePlot(seuratObj, features = c("adt_CD26"), max.cutoff = 'q95', min.cutoff = 'q05') +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("CD8 Prt Exp")

d1c <- FeaturePlot(seuratObj, features = c("adt_CD26"), max.cutoff = 'q95', min.cutoff = 'q05') +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("CD26 Prt Exp")

d1d <- FeaturePlot(seuratObj, features = c("adt_CD161"), max.cutoff = 'q95', min.cutoff = 'q05') +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("CD161 Prt Exp")

d1e <- FeaturePlot(seuratObj, features = c("adt_TCR-Va7.2"), max.cutoff = 'q95', min.cutoff = 'q05') +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("TRAV1-2 Prt Exp")

d1 <- ggarrange(d1a, d1b, d1c, d1d, d1e, ncol = 2, nrow = 3)

```

#Figure 1e
```{r, message = TRUE, fig.show='asis'}
e1a <- FeaturePlot(seuratObj, features = c("TNF")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("TNF Gene Exp")

e1b <- FeaturePlot(seuratObj, features = c("CCL4")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("CCL4 Gene Exp")

e1c <- FeaturePlot(seuratObj, features = c("CCL5")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18))  +
  ggtitle("CCL5 Gene Exp")

e1d <- FeaturePlot(seuratObj, features = c("GNLY")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("GNLY Gene Exp")

e1e <- FeaturePlot(seuratObj, features = c("GZMA")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("GZMA Gene Exp")

e1f <- FeaturePlot(seuratObj, features = c("PRF1")) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(hjust = 0.5, size = 18)) +
  ggtitle("PRF1 Gene Exp")

e1 <- ggarrange(e1a, e1b, e1c, e1d, e1e, e1f, ncol = 2, nrow = 3)


d1 
e1
```

#generate pseudobulk analysis and DEGs between cord blood and adult MR1T
```{r}
pbulk <- CellMembrane::PseudobulkSeurat(seuratObj, groupFields = c("Donor", "cDNA_ID"))

model_matrix <- CellMembrane::DesignModelMatrix(pbulk, contrast_columns = c("Donor"), sampleIdCol = "cDNA_ID")

fit <- CellMembrane::PerformGlmFit(pbulk, model_matrix, assayName = "RNA", test.use = "QLF")

de <- RunPairwiseContrasts(fit, test.use = "QLF")

de_df <- de$`Adult-Cordblood`$differential_expression$table
DE_PBMC <- de$`Adult-Cordblood`$differential_expression$table
DE_PBMC <- de_df[de_df$FDR < 0.2, ]
```

```{r}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Adult"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Cord blood"
```

#Save supplemental table 1
```{r}
write.csv(DE_PBMC, "Supplemental Table 1.csv")
```

#Volcano plot for DEGs of Adult vs Cordblood
```{r, echo = TRUE, fig.show='asis'}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      & DE_PBMC$gene %in% c("GZMK", "GZMA", "GNLY", "GZMM", "PRF1", "GNLY",  "GZMH", "CCR1", "CCR2", "CCR5", "CCR6", "CCL5", "CCL20", "CCL3L3", "CCL3", "CCL4", "CXCR6", "NKG7", "KLRF1", "KLRD1", "KLRC1", "KLRB1", "IL12RB1", "IL12RB2",  "IFNL1", "TNF", "IFNG-AS1", "IFNGR1", "IL18RAP",  "CD244", "CD70", "CD160", "IL26", "CD69", "IL32", "TBX21")
                      ] <- "Adult"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05  
                      & DE_PBMC$gene %in% c("TGFB1", "SELL", "CCR7", "LEF1", "TCF7", "IL13", "SATB1", "CCR9", "TOX")
                      ] <- "Cord blood"

DE_PBMC$delabel <- NA
DE_PBMC$delabel[DE_PBMC$diffexpressed != "NO"] <- DE_PBMC$gene[DE_PBMC$diffexpressed != "NO"]

f1 <- DE_PBMC %>% ggplot(aes(logFC, -log10(FDR), col = diffexpressed, label = delabel)) + geom_point() + theme_minimal() + geom_text_repel(max.overlaps = 50, size = 4) + geom_vline(xintercept=c(-0.5, 0.5), col="dark red") +
    geom_hline(yintercept=-log10(0.05), col="dark red") + scale_color_manual(values = c("dark red", "dark blue", "light grey")) + 
  theme(legend.title = element_blank(), legend.text = element_text(size = 14), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14)) +
  labs(x = "Log Fold Change")
f1
```

#GO pathway analysis of DEGs from adults vs cord blood MR1T
```{r, echo = TRUE, fig.show='asis'}
genes_up <- de_df[
  de_df$logFC >=0.5 & de_df$FDR < 0.05,"gene"]

genes_down <- de_df[
  de_df$logFC <=-0.5 & de_df$FDR < 0.05,"gene"]

#up in adult MR1T
enriched_go <- enrichr(c(genes_up), dbs_go)
enriched_pathway <- enrichr(c(genes_up), dbs_pathway)
g1 <- plotEnrich(enriched_go[[1]], 
                 showTerms = 10, 
                 numChar = 80, 
                 y = "Ratio", 
                 orderBy = "P.value") + 
  theme(axis.text.y = element_text(size = 16), 
        plot.title = element_text(size = 30), 
        axis.title.y = element_blank(), 
        axis.title.x = element_text(size = 14), 
        legend.title = element_text(hjust = 0.5, size = 20), 
        legend.text = element_text(size = 14)) + 
  ggtitle("")

g1
```

#Figure 1f-g
```{r}
ggarrange(f1, g1, ncol = 1, nrow = 2)
```


#Figure 1h-i
```{r, echo = TRUE, fig.show='asis'}
seuratObj <- Seurat::AddModuleScore(seuratObj, features = list(RIRA::GetGeneSet("Cytotoxicity")), name = "MySignature", ctrl = 5)

seuratObj <- UCell::AddModuleScore_UCell(seuratObj, features = list("MySignature" = RIRA::GetGeneSet("Cytotoxicity")))

h1 <- VlnPlot(seuratObj, 
              features = "MySignature1", 
              group.by = "Donor")+
  stat_summary(fun.y = median, 
               geom='point', 
               size = 25, 
               colour = "dark red", 
               shape = 95) + 
  theme(legend.position = 'none',
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 18)) +
  ggtitle("Cytotoxicity")

Idents(seuratObj) = "Donor"

i1 <- FeaturePlot(seuratObj, 
                  features = "MySignature1", 
                  label = TRUE, 
                  label.color = "black", 
                  label.size = 10) + 
  ggtitle("Cytotoxicity") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank()) +
  scale_colour_gradientn(colours = 
                           c("#2E5DAA", "dodgerblue", "lightblue", "gold", "red", "darkred"))


ggarrange(h1, i1, ncol = 2, nrow = 1)

```

#Getting statistics for cytotoxicity score
```{r, echo = TRUE, fig.show='asis'}
ggbetweenstats(
  data  = seuratObj@meta.data[sample(1:nrow(seuratObj@meta.data), 3500),],
  x    = "Donor",
  y    = "Cytotoxicity_UCell",
  title = "Cytotoxicity across Donors"
)
```

__________
#Supplemental Figure 1
```{r}
DefaultAssay(seuratObj) = "ADT"
head(GetAssayData(seuratObj, slot = "data")[, 1:5])

plot(density(GetAssayData(seuratObj, slot = "data")["CD8", ]))
plot(density(GetAssayData(seuratObj, slot = "data")["CD4", ]))
#Double Neg
seuratObj$CD_gate = "neg"
seuratObj$CD_gate[GetAssayData(seuratObj, slot = "data")["CD4", ] < 1.4 & GetAssayData(seuratObj, slot = "data")["CD8", ] < 1.8] = "doubleneg"

seuratObj$CD_gate[GetAssayData(seuratObj, slot = "data")["CD4", ] > 1.4 & GetAssayData(seuratObj, slot = "data")["CD8", ] > 1.8] = "doublepos"

seuratObj$CD_gate[GetAssayData(seuratObj, slot = "data")["CD4", ] < 1.4 & GetAssayData(seuratObj, slot = "data")["CD8", ] > 1.8] = "CD8"

seuratObj$CD_gate[GetAssayData(seuratObj, slot = "data")["CD4", ] > 1.4 & GetAssayData(seuratObj, slot = "data")["CD8", ] < 1.8] = "CD4"

df1 <- data.frame(table(seuratObj$CD_gate, seuratObj$Donor, seuratObj$TRAV))
colnames(df1) = c("CD8", "Donor", "TRAV", "Frequency")

#barplot (not shown in paper)
CD8 <- df1 %>% 
  filter(Donor == "Cordblood") %>% 
  ggplot(aes(x = TRAV, y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", position = "fill") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 24),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  ggtitle("Cord Blood") | 
  df1 %>% 
  filter(Donor == "Adult") %>% 
  ggplot(aes(x = TRAV, y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", position = "fill") +
  ggtitle("Adult")  +
  theme(plot.title = element_text(hjust = 0.5, size = 24),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.title = element_text(size = 16, hjust = 0.5),
        legend.text = element_text(size = 14)) 
CD8

DefaultAssay(seuratObj) = "RNA"

CBTRAV <- df1 %>% 
  filter(Donor == "Cordblood") %>% 
  filter(TRAV == "TRAV1-2") %>% 
  ggplot(aes(x = "", y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood TRAV1-2") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

CBnTRAV <- df1 %>% 
  filter(Donor == "Cordblood") %>% 
  filter(TRAV == "non-TRAV1-2") %>% 
  ggplot(aes(x = "", y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood non-TRAV1-2") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

ATRAV <- df1 %>% 
  filter(Donor == "Adult") %>% 
  filter(TRAV == "TRAV1-2") %>% 
  ggplot(aes(x = "", y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult TRAV1-2") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

AnTRAV <- df1 %>% 
  filter(Donor == "Adult") %>% 
  filter(TRAV == "non-TRAV1-2") %>% 
  ggplot(aes(x = "", y = Frequency, fill = CD8)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult non-TRAV1-2") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

ggarrange(CBTRAV, CBnTRAV, ATRAV, AnTRAV, ncol = 2, nrow = 2)
```

____________________
#Figure 2 - Cord blood TCR Diversity
##Making TRAV1-2,TRAV Column for dual alpha expressing MR1T cells
```{r, echo = TRUE, fig.show='asis'}
meta_CB <- meta %>% 
  filter(Donor == "Cordblood")
p1 <- data.frame(sort(table(meta_CB$TRA_V), decreasing = T))
rownames(p1) <- p1$Var1
p1$TRAV <- p1$Var1
p1$TRAV <- ifelse(p1$TRAV %in% c("TRAV1-1,TRAV1-2", "TRAV1-2,TRAV10", "TRAV1-2,TRAV12-1", "TRAV1-2,TRAV12-2", "TRAV1-2,TRAV12-3", "TRAV1-2,TRAV12-3,TRAV26-2", "TRAV1-2,TRAV13-1", "TRAV1-2,TRAV13-2",  "TRAV1-2,TRAV16", "TRAV1-2,TRAV17", "TRAV1-2,TRAV14/DV4", "TRAV1-2,TRAV14/DV4,TRAV26-2", "TRAV1-2,TRAV19", "TRAV1-2,TRAV2", "TRAV1-2,TRAV20", "TRAV1-2,TRAV21", "TRAV1-2,TRAV23/DV6", "TRAV1-2,TRAV25", "TRAV1-2,TRAV26-1", "TRAV1-2,TRAV26-2", "TRAV1-2,TRAV29/DV5", "TRAV1-2,TRAV3", "TRAV1-2,TRAV39", "TRAV1-2,TRAV4", "TRAV1-2,TRAV5", "TRAV1-2,TRAV6", "TRAV1-2,TRAV8-1", "TRAV1-2,TRAV8-2", "TRAV1-2,TRAV8-3", "TRAV1-2,TRBV8-4", "TRAV1-2,TRAV8-6", "TRAV1-2,TRAV9-2"), "TRAV1-2,TRAV", rownames(p1))


p1 <- p1 %>% 
  group_by(TRAV) %>% 
  summarise(Freq = sum(Freq))
p1 <- p1[order(p1$Freq, decreasing = T), ]

p2 <- p1 %>%
  filter(Freq > 2) 

ga <- p2 %>% 
  ggplot(aes(x = "", y = Freq, fill = reorder(TRAV, -Freq))) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood TRAV") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

meta_A <- meta %>% 
  filter(Donor == "Adult")
p3 <- data.frame(sort(table(meta_A$TRA_V), decreasing = T))

rownames(p3) <- p3$Var1
p3$TRAV <- p3$Var1
p3$TRAV <- ifelse(p3$TRAV %in% c("TRAV1-1,TRAV1-2", "TRAV1-2,TRAV10", "TRAV1-2,TRAV12-1", "TRAV1-2,TRAV12-2", "TRAV1-2,TRAV12-3", "TRAV1-2,TRAV12-3,TRAV26-2", "TRAV1-2,TRAV13-1", "TRAV1-2,TRAV13-2",  "TRAV1-2,TRAV16", "TRAV1-2,TRAV17", "TRAV1-2,TRAV14/DV4", "TRAV1-2,TRAV14/DV4,TRAV26-2", "TRAV1-2,TRAV19", "TRAV1-2,TRAV2", "TRAV1-2,TRAV20", "TRAV1-2,TRAV21", "TRAV1-2,TRAV23/DV6", "TRAV1-2,TRAV25", "TRAV1-2,TRAV26-1", "TRAV1-2,TRAV26-2", "TRAV1-2,TRAV29/DV5", "TRAV1-2,TRAV3", "TRAV1-2,TRAV39", "TRAV1-2,TRAV4", "TRAV1-2,TRAV5", "TRAV1-2,TRAV6", "TRAV1-2,TRAV8-1", "TRAV1-2,TRAV8-2", "TRAV1-2,TRAV8-3", "TRAV1-2,TRBV8-4", "TRAV1-2,TRAV8-6", "TRAV1-2,TRAV9-2"), "TRAV1-2,TRAV", rownames(p3))


p3 <- p3 %>% 
  group_by(TRAV) %>% 
  summarise(Freq = sum(Freq))

p3 <- p3[order(p3$Freq, decreasing = T), ]

p4 <- p3 %>%
  filter(Freq > 2) 

g2 <- p4 %>%
  ggplot(aes(x = "", y = Freq, fill = reorder(TRAV, -Freq))) +
  geom_bar(stat = "identity", width = 0.9) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult TRAV") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30))+
  geom_bar(stat="identity", color='black') +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

ga | g2
```

#TRAJ
```{r, echo = TRUE, fig.show='asis'}
meta_CB <- meta %>% 
  filter(Donor == "Cordblood")
j1 <- data.frame(sort(table(meta_CB$TRA_J), decreasing = T))
j2 <- j1 %>% 
  filter(Freq > 2)

g3 <- j2 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood TRAJ") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30))+
  geom_bar(stat="identity", color='black') +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

meta_A <- meta %>% 
  filter(Donor == "Adult")
j3 <- data.frame(sort(table(meta_A$TRA_J), decreasing = T))
j4 <- j3 %>% 
  filter(Freq > 2)

g4 <- j4 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult TRAJ") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30))+
  geom_bar(stat="identity", color='black') +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

g3 | g4
```

#CDR3a
```{r, echo = TRUE, fig.show='asis'}
meta_CB <- meta %>% 
  filter(Donor == "Cordblood")
ca1 <- data.frame(sort(table(meta_CB$TRA), decreasing = T))
ca2 <- ca1 %>% 
  filter(Freq > 0)

g5 <- ca2 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood CDR3a") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30))+
  geom_bar(stat="identity", color='black') +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

meta_A <- meta %>% 
  filter(Donor == "Adult")
ca3 <- data.frame(sort(table(meta_A$TRA), decreasing = T))
ca4 <- ca3 %>% 
  filter(Freq > 0)

g6 <- ca4 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult CDR3a") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

g5 | g6
```

#TRBV
```{r, echo = TRUE, fig.show='asis'}
meta_CB <- meta %>% 
  filter(Donor == "Cordblood")
b1 <- data.frame(sort(table(meta_CB$TRB_V), decreasing = T))
b2 <- b1 %>% 
  filter(Freq > 2)

g7 <- b2 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Cord Blood TRBV") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30))+
  guides(fill=guide_legend(override.aes=list(colour=NA)))

meta_A <- meta %>% 
  filter(Donor == "Adult")
b3 <- data.frame(sort(table(meta_A$TRB_V), decreasing = T))
b4 <- b3 %>% 
  filter(Freq > 2)

g8 <- b4 %>% 
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 0.9, color = "black") +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Adult TRBV") + 
  theme(plot.title = element_text(hjust = 0.5, size = 30)) +
  guides(fill=guide_legend(override.aes=list(colour=NA)))

g7 | g8
```

#Figure 2a-f
```{r, echo = TRUE, fig.show='asis'}
Figure2 <- ggarrange(ga, g3, g7, g2, g4, g8, ncol = 3, nrow = 2)
Figure2
```

#Figure 2 g-i
##Below code gives tables and then manually copied into prism to graph
```{r}
richness <- function(x){
  
  sum(x > 0)
}

shannon <- function(x){
  
  rabund <- x[x>0]/sum(x)
  -sum(rabund * log(rabund))
}

simpson <- function(x){
  
  n <- sum(x)
  
  sum(x * (x-1)/(x*(n-1)))
}
```

#TRA
```{r, echo = TRUE, results = 'asis'}
#TRA
df <- meta %>% 
  filter(!is.na(TRA)) %>% 
  subset(select = c(SubjectId, TRA))

df1 <- data.frame(table(df$SubjectId, df$TRA))
df1 <- df1 %>% filter(Freq > 0)

df1 %>% 
  group_by(Var1) %>% 
  summarize(sobs = richness(Freq), 
            shannon = shannon(Freq), 
            simpson = simpson(Freq),
            invsimpson = 1/simpson(Freq), 
            n = sum(Freq)) %>% 
  knitr::kable()
```

#TRAJ
```{r, echo = TRUE, results='asis'}
#TRAJ
df <- meta %>% 
  filter(!is.na(TRA_J)) %>% 
  subset(select = c(SubjectId, TRA_J))

df1 <- data.frame(table(df$SubjectId, df$TRA_J))
df1 <- df1 %>% filter(Freq > 0)

df1 %>% 
  group_by(Var1) %>% 
  summarize(sobs = richness(Freq), 
            shannon = shannon(Freq), 
            simpson = simpson(Freq),
            invsimpson = 1/simpson(Freq), 
            n = sum(Freq))
```

#TRBV
```{r, echo = TRUE, results='asis'}
#TRBV
df <- meta %>% 
  filter(!is.na(TRB_V)) %>% 
  subset(select = c(SubjectId, TRB_V))

df1 <- data.frame(table(df$SubjectId, df$TRB_V))
df1 <- df1 %>% filter(Freq > 0)

df1 %>% 
  group_by(Var1) %>% 
  summarize(sobs = richness(Freq), 
            shannon = shannon(Freq), 
            simpson = simpson(Freq),
            invsimpson = 1/simpson(Freq), 
            n = sum(Freq))
```

#TRB
```{r, echo = TRUE, results='asis'}
#TRB
df <- meta %>% 
  filter(!is.na(TRB)) %>% 
  subset(select = c(SubjectId, TRB))

df1 <- data.frame(table(df$SubjectId, df$TRB))
df1 <- df1 %>% filter(Freq > 0)

df1 %>% 
  group_by(Var1) %>% 
  summarize(sobs = richness(Freq), 
            shannon = shannon(Freq), 
            simpson = simpson(Freq),
            invsimpson = 1/simpson(Freq), 
            n = sum(Freq))
```

#TRBJ
```{r, echo = TRUE, results = 'asis'}
#TRBJ
df <- meta %>% 
  filter(!is.na(TRB_J)) %>% 
  subset(select = c(SubjectId, TRB_J))

df1 <- data.frame(table(df$SubjectId, df$TRB_J))
df1 <- df1 %>% filter(Freq > 0)

df1 %>% 
  group_by(Var1) %>% 
  summarize(sobs = richness(Freq), 
            shannon = shannon(Freq), 
            simpson = simpson(Freq),
            invsimpson = 1/simpson(Freq), 
            n = sum(Freq))
```

#Supplemental Figure 3 - TRAV Surface Expression
```{r, echo = TRUE, fig.show='asis'}
seuratObjCB = subset(seuratObj, Donor == "Cordblood")
seuratObjCBTCR = subset(seuratObjCB, TRAV %in% c("non-TRAV1-2", "TRAV1-2"))

Supplemental3 <- VlnPlot(seuratObjCBTCR, features = "adt_TCR-Va7.2", group.by = "TRAV") + 
  stat_summary(fun.y = median, geom='point', size = 25, colour = "dark red", shape = 95) +
  theme(legend.position = 'none') + 
  ggtitle("Cord Blood MR1T Cells Surface Protein Expression of TRAV1-2") +
  theme(axis.title.x = element_blank())

Supplemental3
```

#Supplemental - not shown. TRAJ usage of TRAV1-2 Positive vs Negative
```{r}
seuratObjCBTRAVneg = subset(seuratObjTRAVneg, Donor == "Cordblood")
seuratObjATRAVneg = subset(seuratObjTRAVneg, Donor == "Adult")

seuratObjCBTRAVpos = subset(seuratObjTRAVpos, Donor == "Cordblood")
seuratObjATRAVpos = subset(seuratObjTRAVpos, Donor == "Adult")

Plot_bar_topvalues(so=seuratObjCBTRAVpos, feat = "TRA_J", title = "CB TRAV1-2+ TRAJ Usage", thr = 1) | Plot_bar_topvalues(so=seuratObjCBTRAVneg, feat = "TRA_J", title = "CB TRAV1-2- TRAJ Usage", thr = 1)
```

___________________
#Figure 3D - DEGs of TRAV1-2+ vs non-TRAV1-2
```{r}
Idents(seuratObjCBTCR) = "TRAV"
DE_PBMC  = FindMarkers(seuratObjCBTCR, ident.1 = c("TRAV1-2"), ident.2 = c("non-TRAV1-2"), logfc.threshold = 0.5, test.use = "wilcox", min.pct = 0, min.diff.pct = 0, min.cells.feature = 2, min.cells.group = 2)

de_df <- DE_PBMC
DE_PBMC <- tibble::rownames_to_column(DE_PBMC, "gene")
```

#GO pathways for DEGs between TRAV1-2+ and non-TRAV1-2 (not included in paper)
```{r, echo = TRUE, fig.show='asis'}
de_df$gene <- rownames(de_df)

genes_up <- de_df[
  de_df$avg_log2FC >=0.5 & de_df$p_val_adj < 0.05,"gene"]

genes_down <- de_df[
  de_df$avg_log2FC <=-0.5 & de_df$p_val_adj < 0.05,"gene"]

enriched_go <- enrichr(c(genes_up), dbs_go)
enriched_pathway <- enrichr(c(genes_up), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated Pathways in Cord blood TRAV1-2+ MR1T Cells")

p1 <- plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated Pathways in Cord blood TRAV1-2+ MR1T Cells")

enriched_go <- enrichr(c(genes_down), dbs_go)
enriched_pathway <- enrichr(c(genes_down), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated Pathways in Cord blood TRAV1-2- MR1T Cells")

p2 <- plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated Pathways in Cord Blood TRAV1-2- MR1T Cells")

```

#Generating supplemental table 2
```{r}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$avg_log2FC > 0.5 & DE_PBMC$p_val_adj < 0.05 
                      ] <- "Cord blood TRAV1-2+ MR1T Cells"
DE_PBMC$diffexpressed[DE_PBMC$avg_log2FC < -0.5 & DE_PBMC$p_val_adj < 0.05 
                        ] <- "Cord Blood TRAV1-2- MR1T Cells"

write.csv(DE_PBMC, "Supplemental Table 2.csv")
```

#Figure 3D - volcano plot of DEGs between TRAV1-2 and non-TRAV1-2 MR1Ts
```{r, echo = TRUE, fig.show='asis'}
#Volcano Plot
DE_PBMC$diffexpressed <- ""
DE_PBMC$diffexpressed[DE_PBMC$avg_log2FC > 0.5 & DE_PBMC$p_val_adj < 0.05 
                      & DE_PBMC$gene %in% c("TRAV1-2", "KLRB1", "NCR3", "TRBV6-1", "TRBV6-2", "TRBV6-4", "LMNA", "ANXA1", "GNA15", "SH2D2A", "RUNX3", "VIM", "PRNP", "KLRK1", "CSRNP1", "JUND", "FLNA", "ISG15", "ZBTB16", "FOS", "ZYX")
                      ] <- "Cord blood TRAV1-2+ MR1T Cells"
DE_PBMC$diffexpressed[DE_PBMC$avg_log2FC < -0.5 & DE_PBMC$p_val_adj < 0.05 
                      #& DE_PBMC$gene %in% c()
                        ] <- "Cord Blood TRAV1-2- MR1T Cells"

DE_PBMC$delabel <- NA
DE_PBMC$delabel[DE_PBMC$diffexpressed != ""] <- DE_PBMC$gene[DE_PBMC$diffexpressed != ""]

Fig3D <- DE_PBMC %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), col = diffexpressed, label = delabel)) + geom_point() + theme_minimal() + geom_text_repel(max.overlaps = 50, size = 4) + geom_vline(xintercept=c(-0.5, 0.5), col="dark red") +
    geom_hline(yintercept=-log10(0.05), col="dark red") + scale_color_manual(values = c("light grey", "dark blue", "dark red"))+ggtitle("Differential Expression of Cord Blood TRAV1-2 Positive vs Negative MR1T Cells") +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 18), 
        axis.title.y = element_text(size = 18), 
        legend.text = element_text(size = 14), 
        plot.title = element_text(hjust = 0.5, size = 22)) 
Fig3D
```

________________
#Figure 5 - Difference in TRAV1-2+ Adult vs Cord blood 
```{r, echo = TRUE, fig.show='asis'}
seuratObjTRAVpos <- 
  Seurat::AddModuleScore(seuratObjTRAVpos, 
                                           features = 
                           list(RIRA::GetGeneSet("Cytotoxicity")), 
                         name = "MySignature", ctrl = 5)

seuratObjTRAVpos <- UCell::AddModuleScore_UCell(seuratObjTRAVpos, 
                                                features = 
                                                  list("MySignature" = 
                                                      RIRA::GetGeneSet("Cytotoxicity")))

Sup5a <- VlnPlot(seuratObjTRAVpos, 
                 features = "MySignature1", 
                 group.by = "Donor")+
  stat_summary(fun.y = median, 
               geom='point', 
               size = 25,
               colour = "dark red",
               shape = 95)+ 
  theme(legend.position = 'none', 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 14)) +
  ggtitle("Cytotoxicity TRAV1-2+ Cells")

Idents(seuratObjTRAVpos) = "Donor"
Sup5b <- FeaturePlot(seuratObjTRAVpos, 
                     features = "MySignature1", 
                     label = TRUE, 
                     label.color = "black", 
                     label.size = 10) + 
  ggtitle("Cytotoxicity TRAV1-2+ Cells") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  scale_colour_gradientn(colours = 
                           c("#2E5DAA", "dodgerblue", "lightblue", "gold", "red", "darkred"))

Sup5a | Sup5b
```

#generate statistics for cytotoxicity score
```{r, echo = TRUE, fig.show='asis'}
ggbetweenstats(
  data  = seuratObjTRAVpos@meta.data[sample(1:nrow(seuratObjTRAVpos@meta.data), 1100),],
  x    = "Donor",
  y    = "Cytotoxicity_UCell",
  title = "Cytotoxicity across Donors TRAV1-2+"
)
```

#Additional TRAV1-2 expression plots, not included in paper
```{r}
meta1 <- meta %>% 
    filter(!is.na(TRA) & !is.na(TRB)) %>% 
  filter(Donor == c("Adult"))
meta1$TR_AB <- paste0(meta1$TRA, "__", meta1$TRB)
p1 <- data.frame(sort(table(meta1$TR_AB), decreasing = T))


meta2 <- meta %>% 
    filter(!is.na(TRA) & !is.na(TRB)) %>% 
  filter(Donor == "Cordblood")
meta2$TR_AB <- paste0(meta2$TRA, "__", meta2$TRB)
p2 <- data.frame(sort(table(meta2$TR_AB), decreasing = T))


p3 <- p1 %>%  filter(Freq > 2) %>% ggplot(aes(x=Var1, y = Freq))+geom_col()+theme(axis.text.x=element_text(angle = 45, vjust=1, hjust=1), plot.title = element_text(hjust = 0.5, size = 24))+ggtitle("Adult")

p4 <- p2 %>%  filter(Freq > 1) %>% ggplot(aes(x=Var1, y = Freq))+geom_col()+theme(axis.text.x=element_text(angle = 45, vjust=1, hjust=1), plot.title = element_text(hjust = 0.5, size = 24))+ggtitle("Cord Blood")

p3 | p4 + facet_wrap(~Var1, nrow = 2, scales = "free_y")

meta3 <- meta %>% 
  filter(Donor == "Cordblood")
p5 <- data.frame(sort(table(meta3$TRAV), decreasing = T))

meta4 <- meta %>% 
  filter(Donor == "Adult")
p6 <- data.frame(sort(table(meta4$TRAV), decreasing = T))

p7 <- p5 %>%  filter(Freq > 2) %>% ggplot(aes(x=Var1, y = Freq))+geom_col()+theme(axis.text.x=element_text(angle = 45, vjust=1, hjust=1), plot.title = element_text(hjust = 0.5, size = 24))+ggtitle("Cord Blood")

p8 <- p6 %>%  filter(Freq > 2) %>% ggplot(aes(x=Var1, y = Freq))+geom_col()+theme(axis.text.x=element_text(angle = 45, vjust=1, hjust=1), plot.title = element_text(hjust = 0.5, size = 24))+ggtitle("Adult")

p7 | p8

ggarrange(p7, p8, ncol = 1, nrow = 2)
```

______________
##Clone Analysis - second half of paper
#Save clone analysis if changes
```{r, fig.show='asis'}
#saveRDS(seuratObj2, file = "./CB_clone_seurat.rds")
```

#Load single-cell MR1T clone seuratObj
```{r}
fn <- 'CB_clone_seurat.rds'
seuratObj2 <- readRDS(fn)
```

#Figure 7A-F
```{r, echo = TRUE, fig.show='asis'}
Idents(seuratObj2) = "StimulationCellType"
my_levels <- c("Adult Unstim", "Adult Stim", "CB Unstim", "CB Stim")
seuratObj2@meta.data$StimulationCellType <- factor(x = seuratObj2@meta.data$StimulationCellType, levels = c("Adult Unstim", "Adult Stim", "CB Unstim", "CB Stim"))

IFNG <- VlnPlot(seuratObj2, features = c("IFNG"), split.by = "StimulationCellType") + stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

TNF <- VlnPlot(seuratObj2, features = c("TNF"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

GZMB <- VlnPlot(seuratObj2, features = c("GZMB"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CCL3 <- VlnPlot(seuratObj2, features = c("CCL3"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CCL4 <- VlnPlot(seuratObj2, features = c("CCL4"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CSF2 <- VlnPlot(seuratObj2, features = c("CSF2"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

Fig5 <- ggarrange(IFNG, TNF, GZMB, CCL3, CCL4, CSF2)
Fig5
```

#Supplemental Figure 6
```{r}
Idents(seuratObj2) = "SubjectId"

IFNG <- VlnPlot(seuratObj2, features = c("IFNG"), split.by = "StimulationCellType") + stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

TNF <- VlnPlot(seuratObj2, features = c("TNF"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

GZMB <- VlnPlot(seuratObj2, features = c("GZMB"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CCL3 <- VlnPlot(seuratObj2, features = c("CCL3"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CCL4 <- VlnPlot(seuratObj2, features = c("CCL4"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

CSF2 <- VlnPlot(seuratObj2, features = c("CSF2"), split.by = "StimulationCellType") +
  stat_summary(fun.y = mean, geom='point', size = 25, colour = "dark red", shape = 95) + theme(legend.position = 'none') + 
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 16),
        plot.title = element_text(size = 20))

Supp_Clone <- ggarrange(IFNG, TNF, GZMB, CCL3, CCL4, CSF2)
Supp_Clone
```


#Figure 7B and Supplemental Figure 7 pseudobulk analysis
```{r}
pbulk <- CellMembrane::PseudobulkSeurat(seuratObj2, groupFields = c("StimulationCellType", "cDNA_ID"))

model_matrix <- CellMembrane::DesignModelMatrix(pbulk, contrast_columns = c("StimulationCellType"), sampleIdCol = "cDNA_ID")

fit <- CellMembrane::PerformGlmFit(pbulk, model_matrix, assayName = "RNA", test.use = "QLF")

de <- RunPairwiseContrasts(fit, test.use = "QLF")
```

```{r}
de_df <- de$`Adult.Unstim-CB.Unstim`$differential_expression$table
DE_PBMC <- de$`Adult.Unstim-CB.Unstim`$differential_expression$table
DE_PBMC <- de_df[de_df$FDR < 0.2, ]
```

#generating supplemental table 3
```{r}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Adult Unstimulated"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Cordblood Unstimulated"

table(DE_PBMC$diffexpressed)
```

```{r}
write.csv(DE_PBMC, "Supplemental Table 3.csv")
```

##Pathway analysis (not shown in paper)
```{r, echo = TRUE, fig.show='asis'}
genes_up <- DE_PBMC[
  DE_PBMC$logFC >=0.5 & DE_PBMC$FDR < 0.05,"gene"]

genes_down <- DE_PBMC[
  DE_PBMC$logFC <=-0.5 & DE_PBMC$FDR < 0.05,"gene"]

#Run to look at genes up in adult unstim
enriched_go <- enrichr(c(genes_up), dbs_go)
enriched_pathway <- enrichr(c(genes_up), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated in Adult Unstimulated MR1T Cells")

#Run to look at genes up in cord blood unstim
enriched_go <- enrichr(c(genes_down), dbs_go)
enriched_pathway <- enrichr(c(genes_down), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated in Cord Blood Unstimulated MR1T Cells")
```


##Volcano plot for DE of Unstimulated Adult vs Cordblood
```{r, echo = TRUE, fig.show='asis'}
DE_PBMC$diffexpressed <- ""
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      & DE_PBMC$gene %in% c("GZMH", "PRF1", "CD8A", "CD8B2", "CCR5", "CXCR6", "CXCR3", "LTB", "FASLG", "ENO1", "ENO2", "ALDOC", "GAPDH", "PFKP", "PGAM5", "PGK1", "PKM", "CEBPD")
                      ] <- "Adult Unstimulated"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05 
                      & DE_PBMC$gene %in% c("SELL", "CCR7", "IL7R", "FUOM", "LEF1", "RGS10", "CCL3", "CCL3L3", "CCL4", "CCL4L2", "CCL23", "CCL18", "CCL1", "CCL22", "CCL24", "CCL8", "CCL13", "CCL7", "CCL2", "CXCL9", "CXCL12", "CXCL3", "CXCL16", "IFI30", "IFI27", "IFI44L", "IFIT3", "IFIT1", "IFI6", "IFI27L1", "IFI27L2", "ISG15", "MX2", "IRF4", "IRF9", "IRF8", "IRF1-AS1", "HERC5", "IFIT1", "IFIT3", "IL17C")
                      ] <- "Cord Blood Unstimulated"

DE_PBMC$delabel <- NA
DE_PBMC$delabel[DE_PBMC$diffexpressed != ""] <- DE_PBMC$gene[DE_PBMC$diffexpressed != ""]

a4 <- DE_PBMC %>% ggplot(aes(logFC, -log10(FDR), col = diffexpressed, label = delabel)) + geom_point() + theme_minimal() + geom_text_repel(max.overlaps = 50, size = 4) + geom_vline(xintercept=c(-0.5, 0.5), col="dark red") +
    geom_hline(yintercept=-log10(0.05), col="dark red") + scale_color_manual(values = c("light grey", "dark red", "dark blue"))+ggtitle("Unstimulated") + theme(legend.position = "none") + theme(plot.title = element_text(hjust = 0.5, size = 20), axis.title.x = element_text(size = 14),  axis.title.y = element_text(size = 14), legend.text = element_text(size = 14), axis.text = element_text(size = 12))  +
  labs(x = "Log Fold Change")
a4
```

#Adult Stim vs Cord Blood Stim
```{r}
de_df <- de$`Adult.Stim-CB.Stim`$differential_expression$table
DE_PBMC <- de$`Adult.Stim-CB.Stim`$differential_expression$table
DE_PBMC <- de_df[de_df$FDR < 0.2, ]
```

##Pathway analysis (not shown in paper)
```{r, echo = TRUE, fig.show='asis'}
genes_up <- DE_PBMC[
  DE_PBMC$logFC >=0.5 & DE_PBMC$FDR < 0.05,"gene"]

genes_down <- DE_PBMC[
  DE_PBMC$logFC <=-0.5 & DE_PBMC$FDR < 0.05,"gene"]

#Run to look at genes up in adult stim
enriched_go <- enrichr(c(genes_up), dbs_go)
enriched_pathway <- enrichr(c(genes_up), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated in Adult Stimulated MR1T Cells")

#Run to look at genes up in cord blood stim
enriched_go <- enrichr(c(genes_down), dbs_go)
enriched_pathway <- enrichr(c(genes_down), dbs_pathway)
plotEnrich(enriched_go[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value") + ggtitle("Upregulated in Cord Blood Stimulated MR1T Cells")
```


```{r}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Adult"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05 
                      ] <- "Cord Blood"

table(DE_PBMC$diffexpressed)
```

#Save supplemental table 4
```{r}
write.csv(DE_PBMC, "Supplemental Table 4.csv")
```


##Volcano plot for DE of Stimulated Adult vs Cordblood
```{r, echo = TRUE, fig.show='asis'}
DE_PBMC$diffexpressed <- "NO"
DE_PBMC$diffexpressed[DE_PBMC$logFC > 0.5 & DE_PBMC$FDR < 0.05 
                      & DE_PBMC$gene %in% c("IL26", "GZMH", "GZMM", "GNLY", "LTB", "RGCC", "TNFSF14", "NR4A3", "FASLG", "TNF", "IL32", "IL17RA", "ALDOC", "ENO2", "HK2", "PGAM1", "PGK1", "PKM", "PFKP", "CTLA4", "HAVCR2", "LAG3", "CD244", "TNFRSF14", "CD160", "CST7", "CD8A", "CCL28", "CCL5", "CXCR6", "GZMM", "GZMH", "GNLY")
                      ] <- "Adult"
DE_PBMC$diffexpressed[DE_PBMC$logFC < -0.5 & DE_PBMC$FDR < 0.05 
                      & DE_PBMC$gene %in% c("IFI30", "IFI27", "IFI6", "IFIT3", "ISG15", "IRF9", "IRF8", "MX1", "IL7R", "CCR7", "CCL1", "CCL7", "CCL2", "CXCL9", "CXCL2", "CXCL3", "CXCL16")
                      ] <- "Cordblood"

DE_PBMC$delabel <- NA
DE_PBMC$delabel[DE_PBMC$diffexpressed != "NO"] <- DE_PBMC$gene[DE_PBMC$diffexpressed != "NO"]

b4 <- DE_PBMC %>% ggplot(aes(logFC, -log10(FDR), col = diffexpressed, label = delabel)) + geom_point() + theme_minimal() + geom_text_repel(max.overlaps = 50, size = 4) + geom_vline(xintercept=c(-0.5, 0.5), col="dark red") +
    geom_hline(yintercept=-log10(0.05), col="dark red") + scale_color_manual(values = c("dark red", "dark blue", "light grey"))+ggtitle("Stimulated") + theme(plot.title = element_text(hjust = 0.5, size = 20), axis.title.x = element_text(size = 14),  axis.title.y = element_text(size = 14), legend.text = element_text(size = 14), legend.title = element_blank(), axis.text = element_text(size = 12)) +
  labs(x = "Log Fold Change")

Fig5GH <- a4 | b4
Fig5GH
```


sessionInfo()
R version 4.4.1 (2024-06-14)
Platform: x86_64-apple-darwin20
Running under: macOS Sonoma 14.7.8

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-x86_64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Toronto
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] ggstatsplot_0.12.4 forcats_1.0.0      readxl_1.4.5       enrichR_3.2       
 [5] ggpubr_0.6.0       tibble_3.3.0       ggrepel_0.9.6      dplyr_1.1.4       
 [9] ggplot2_3.5.2      UCell_2.13.1       Rlabkey_3.3.0      jsonlite_2.0.0    
[13] httr_1.4.7         Rdiscvr_1.0.0      Seurat_5.3.0       SeuratObject_5.1.0
[17] sp_2.2-0           RIRA_1.0.0         CellMembrane_1.0.0

loaded via a namespace (and not attached):
  [1] matrixStats_1.5.0           spatstat.sparse_3.1-0       insight_0.20.4             
  [4] RColorBrewer_1.1-3          prabclus_2.3-4              tools_4.4.1                
  [7] mlr3learners_0.12.0         sctransform_0.4.2           backports_1.5.0            
 [10] R6_2.6.1                    statsExpressions_1.5.5      HDF5Array_1.32.1           
 [13] lazyeval_0.2.2              uwot_0.2.3                  mlr3data_0.9.0             
 [16] rhdf5filters_1.16.0         withr_3.0.2                 mlr3viz_0.10.1             
 [19] gridExtra_2.3               progressr_0.15.1            cli_3.6.5                  
 [22] Biobase_2.64.0              mlr3inferr_0.1.0            mlr3hyperband_1.0.0        
 [25] spatstat.explore_3.5-2      fastDummies_1.7.5           diptest_0.77-1             
 [28] robustbase_0.99-4-1         spatstat.data_3.1-6         mlr3tuningspaces_0.6.0     
 [31] ggridges_0.5.6              pbapply_1.7-4               mlr3tuning_1.4.0           
 [34] R.utils_2.13.0              paradox_1.0.1               WriteXLS_6.7.0             
 [37] parallelly_1.45.1           limma_3.60.6                rstudioapi_0.17.1          
 [40] generics_0.1.4              ica_1.0-3                   spatstat.random_3.4-1      
 [43] car_3.1-2                   Matrix_1.7-0                S4Vectors_0.42.1           
 [46] abind_1.4-8                 R.methodsS3_1.8.2           lifecycle_1.0.4            
 [49] yaml_2.3.10                 edgeR_4.2.2                 carData_3.0-5              
 [52] SummarizedExperiment_1.34.0 rhdf5_2.48.0                SparseArray_1.4.8          
 [55] Rtsne_0.17                  paletteer_1.6.0             grid_4.4.1                 
 [58] promises_1.3.3              dqrng_0.4.1                 crayon_1.5.3               
 [61] miniUI_0.1.2                lattice_0.22-6              beachmat_2.20.0            
 [64] cowplot_1.2.0               scGate_1.7.2                zeallot_0.1.0              
 [67] pillar_1.11.0               knitr_1.50                  GenomicRanges_1.56.2       
 [70] rjson_0.2.23                fpc_2.2-13                  future.apply_1.20.0        
 [73] codetools_0.2-20            glue_1.8.0                  spacefillr_0.4.0           
 [76] spatstat.univar_3.1-4       data.table_1.17.8           vctrs_0.6.5                
 [79] png_0.1-8                   spam_2.11-1                 cellranger_1.1.0           
 [82] gtable_0.3.6                rematch2_2.1.2              datawizard_0.12.3          
 [85] kernlab_0.9-33              xfun_0.52                   S4Arrays_1.4.1             
 [88] mime_0.13                   DropletUtils_1.24.0         correlation_0.8.5          
 [91] NBZIMM_1.0                  survival_3.7-0              SingleCellExperiment_1.26.0
 [94] statmod_1.5.0               fitdistrplus_1.2-4          ROCR_1.0-11                
 [97] nlme_3.1-166                naturalsort_0.1.3           bbotk_1.6.0                
[100] RcppAnnoy_0.0.22            GenomeInfoDb_1.40.1         mlr3pipelines_0.9.0        
[103] mlr3_1.1.0                  irlba_2.3.5.1               KernSmooth_2.23-24         
[106] mlr3mbo_0.3.0               mlr3measures_1.0.0          colorspace_2.1-1           
[109] BiocGenerics_0.50.0         nnet_7.3-19                 mlr3misc_0.18.0            
[112] DESeq2_1.44.0               tidyselect_1.2.1            curl_6.4.0                 
[115] compiler_4.4.1              BiocNeighbors_1.22.0        lgr_0.5.0                  
[118] DelayedArray_0.30.1         plotly_4.11.0               bayestestR_0.14.0          
[121] checkmate_2.3.2             scales_1.4.0                DEoptimR_1.1-4             
[124] lmtest_0.9-40               palmerpenguins_0.1.1        stringr_1.5.1              
[127] digest_0.6.37               goftest_1.2-3               spatstat.utils_3.1-5       
[130] rmarkdown_2.29              XVector_0.44.0              htmltools_0.5.8.1          
[133] pkgconfig_2.0.3             SingleR_2.6.0               sparseMatrixStats_1.16.0   
[136] MatrixGenerics_1.16.0       fastmap_1.2.0               rlang_1.1.6                
[139] htmlwidgets_1.6.4           UCSC.utils_1.0.0            shiny_1.11.1               
[142] DelayedMatrixStats_1.26.0   farver_2.1.2                zoo_1.8-14                 
[145] BiocParallel_1.38.0         mclust_6.1.1                R.oo_1.27.1                
[148] BiocSingular_1.20.0         magrittr_2.0.3              modeltools_0.2-24          
[151] scuttle_1.14.0              GenomeInfoDbData_1.2.12     dotCall64_1.2              
[154] parameters_0.22.1           patchwork_1.3.1             Rhdf5lib_1.26.0            
[157] mlr3fselect_1.4.0           Rcpp_1.1.0                  reticulate_1.43.0          
[160] stringi_1.8.7               zlibbioc_1.50.0             MASS_7.3-61                
[163] plyr_1.8.9                  flexmix_2.3-20              MAST_1.30.0                
[166] parallel_4.4.1              listenv_0.9.1               deldir_2.0-4               
[169] mlr3verse_0.3.1             splines_4.4.1               tensor_1.5.1               
[172] locfit_1.5-9.12             mlr3cluster_0.1.11          mlr3filters_0.8.1          
[175] ranger_0.17.0               igraph_2.1.4                uuid_1.2-1                 
[178] spatstat.geom_3.5-0         ggsignif_0.6.4              effectsize_0.8.9           
[181] RcppHNSW_0.6.0              reshape2_1.4.4              stats4_4.4.1               
[184] ScaledMatrix_1.12.0         evaluate_1.0.4              httpuv_1.6.16              
[187] RANN_2.6.2                  tidyr_1.3.1                 purrr_1.1.0                
[190] polyclip_1.10-7             future_1.67.0               clue_0.3-66                
[193] scattermore_1.2             rsvd_1.0.5                  broom_1.0.9                
[196] xtable_1.8-4                RSpectra_0.16-2             rstatix_0.7.2              
[199] later_1.4.2                 viridisLite_0.4.2           class_7.3-22               
[202] IRanges_2.38.1              cluster_2.1.6               globals_0.18.0     
